troubleshooting

### 2024-05-02 README.Rmd
cleared all notes except for global binding ones
changed github actions

<!-- badges: start -->
<!-- badges: end -->

@@

**Making mspms generically useful**.   
1. *prepare_peaks()*:Takes two input files from PEAKS and combines them. 
2. *prepare_pd()*: prepares exported files from proteome discoverer.  
  - *extract_design_matrix_pd()*: Extracts the design matrix from the proteome discoverer file.  
3. *prepare_spectronaut()*: prepares exported files from spectronaut - note in development, not yet implemented.  
4. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.

**Data Proessing/ Normalization**. 
1. *normalyze()*: Normalizes values and then does a reverse log2 transformation.    
2. *handle_outliers()*: Looks for outliers across replicates. Removes them.     
3. *impute()*: Imputes data for missing values (not including NAs introduced by handle_outliers()).    
4. *join_with_library()*: Joins the normalyzed data with the peptide library sequences.  
5. *add_cleavages()*: Figures out the locations of the detected clevages within the library of peptide sequences used. Is it cleaved at the N or C terminus, or both?  Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.  
6. *polish()*: combines the cleavage info into one column. Discards peptides that were cleaved on both sides or not at all.  
7. *prepare_for_stats()*: Reshapes the data into a long format and appends the data in the design matrix to make statistical testing easy to perform.

@@

**Data Visualization**.   
1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an interactive heatmap to visualize overall patterns in the data.    
2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
3. *plot_time_course()*: Plot peptides over time by condition.  
4. *plot_cleavage_motif()*: Visualize the sequence specificity of the cleavage sites. Designed to be similar to what is implemented in IceLogo.

@@

@@

We might want to calculate all possible cleavages for the peptide library sequences. This is useful for downstream analysis, especially when we are looking at the specificity of the cleavage sites via  as this requires a background of all possible cleavages

@@

msp-ms uses the normalyzede package to do normalization under the hood.

Now we can normalyze the data.
```{r}
normalyzed_data = normalyze(peaks_prepared_data,design_matrix)

@@

We have a lot of missing, or 0 values. For these, we need to impute them so we can do downstream statistics
Data is imputated by ....

@@

Next we need to join everything with the sequences of the peptide library
```{r}
joined_with_library = join_with_library(imputed)

@@

### Calcuating clevages.

@@

We can also calculate the log 2 fc. The comparisons here are the same as for the T tests
```{r}
# calclulate log2fc
log2fc = mspms::mspms_log2fc(prepared_for_stats)

@@

Now that we have this data we can visualize these results easily using the ggplot2 package.

@@

![](www/plot_heatmap_output.png)

>>>>

<!-- badges: start -->
[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

@@

**Making mspms generically useful**.   
1. *prepare_peaks()*:Takes two input files from PEAKS and combines them. 
2. *prepare_pd()*: prepares exported files from proteome discoverer.  
3. *extract_design_matrix_pd()*: Extracts the design matrix from the proteome discoverer file.  
4. *prepare_spectronaut()*: prepares exported files from spectronaut - note in development, not yet implemented.  
5. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.

**Data Proessing/ Normalization**.  
1. *normalyze()*: Normalizes values.     
2. *handle_outliers()*: Looks for outliers across replicates. Removes them.       
3. *impute()*: Imputes data for missing values (not including NAs introduced by handle_outliers()).    
4. *join_with_library()*: Joins the normalyzed data with the peptide library sequences.  
5. *add_cleavages()*: Figures out the locations of the detected clevages within the library of peptide sequences used. Is it cleaved at the N or C terminus, or both?  Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.    
6. *polish()*: combines the cleavage info into one column. Discards peptides that were cleaved on both sides or not at all.  
7. *prepare_for_stats()*: Reshapes the data into a long format and appends the data in the design matrix to make statistical testing easy to perform.

@@

**Data Visualization**.   
1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an interactive heatmap to visualize overall patterns in the data.    
2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
3. *plot_time_course()*: Plot peptides over time by condition.  
4. *plot_cleavage_motif()*: Visualize the sequence specificity of the cleavage sites. Designed to be equivalent to what is implemented in IceLogo.

@@

### Formating Proteome Discoverer File Outputs.

We can prepare proteome discover files for analysis with mspms as follows.

```{r}
prepared_proteome_discoverer = prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
```

We can also extract a design matrix from the naming scheme used by proteome discoverer as follows.

```{r}
 design_matrix = extract_design_matrix_pd("tests/testdata/proteome_discoverer_output.xlsx")
```

@@

We might want to calculate all possible cleavages for the peptide library sequences. This is useful for downstream analysis, especially when we are looking at the specificity of the cleavage sites via plot_cleavage_motif() as this requires a background of all possible cleavages

@@

mspms uses the NormalyzerDE package to do normalization under the hood.

We can normalyze the data as follows:  
```{r}
normalyzed_data = normalyze(peaks_prepared_data,design_matrix)

@@

We have a lot of missing, or 0 values. For these, we need to impute them so we can do downstream statistics.

@@

Next we need to join everything with the sequences of the peptide library.  
```{r}
joined_with_library = join_with_library(imputed)

@@

### Calcuating cleavages.

@@

We can also calculate the log 2 fold change. The comparisons here are the same as for the T tests.  
```{r}
# calclulate log2fc
log2fc = mspms::mspms_log2fc(prepared_for_stats)

@@

Now that we have this data we can visualize these results easily using the ggplot2 package as follows.

@@

![](plot_heatmap_output.png)
